---
title: "AML Assignment 1"
author: "Yiting Wang"
date: "2018-02-20"
output: html_document
---

## Question 2

```{r}
# load data
data <- read.csv("C:/Tahoe_Healthcare_Data.csv")
View(data)
```

### (i)
```{r}
readmit <- sum(data$readmit30)
readmit
unit_loss <- 8000
total_loss <- readmit * unit_loss
total_loss
```

We take the dataset as representative of what will happen in a given year if nothing is done to reduce the readmissions rate. From the dataset, we can calculate the total number of re-admitted patients, which is 998. We mutiply this number by the loss per re-admitted patient \$8,000, and get the total lost of $7,984,000.

### (ii)
```{r}
program_unit_cost <- 1200
patients <- nrow(data)
patients 
program_total_cost <- patients * program_unit_cost
program_total_cost

reduce_rate <- 0.4
saving <- total_loss * reduce_rate
saving

net_change <- program_total_cost - saving
net_change
```

From the dataset we know there are total 4,382 patients in a given year. And the cost of CareTracker program is \$1,200 per patient. So if the hospital implements CareTracker for all patients, the total cost will be $5,258,400, which is a quite high figure. 

And we know CareTracker reduced the incidence of readmissions by 40% in the pilot study, which means it can save 40% of the total loss we calculated in problem (i). So the hospital can save $3,193,600. (40% * 998 * 8000)

The net change in cost is \$2,064,800. If the hospital implements CareTracker for each patient, the cost compared to status quo, will increase $2,064,800, which means the money hospital spends on implementing CareTracker is higher than the money it can save from CareTracker's results. 

From an economic point of view, the hospital should not implement CareTracker for all AMI patients. Maybe it should choose some particular patients to implement CareTracker.

### (iii)
```{r}
best_saving <- reduce_rate * readmit * unit_loss - readmit * program_unit_cost 
best_saving
```

Suppose the hospital can find all the 998 re-admitted patients correctly, and implement CareTracker to them (cost is \$1,200 per patient), which will cost \$1,197,600. However, 40% of the 998 re-admitted patients will be free from readmission, and each will save \$8,000, so the total saving is \$3,193,600. So the upper bound on the possible savings should be $1,996,000. The hospital cannot expect to make more savings by implementing CareTracker.

### (iv)
```{r}
library(dplyr)

outcome = matrix(NA, nrow = 76, ncol = 2)

for (s in 25:100) {
  data_new <- data %>% mutate(estimate = ifelse(severity.score >= s, 1, 0))
  cost <- 1200 * sum(data_new$estimate)
  save <- (sum(data_new$readmit30 + data_new$estimate == 2)) * 0.4 * 8000
  net <- save - cost
  outcome[s - 24, 1] = s
  outcome[s - 24, 2] = net
}

plot(outcome[,1], outcome[,2], type = "l", xlab = "S*", ylab = "savings")

max_saving <- max(outcome[,2])
max_saving

best_s <- outcome[,1][which(outcome[,2] == max_saving)]
best_s
```

From the graph and outcome matrix, we can find out that the best value for the threshold S* is 42, under this circumstance, hospital can obtain a cost savings of $136,800 over the status quo.


### (v)
```{r}
glm.fit <- glm(readmit30 ~ age + female + flu_season + ed_admit + severity.score + comorbidity.score, data = data, family = binomial)

summary(glm.fit)

```


### (vi)
```{r}
data <- data %>% mutate(fit = predict(glm.fit, data, type = "response"))

outcome2 <- matrix(NA, nrow = 81, ncol = 2)

n = 0

for (p in seq(0.1, 0.9, by = 0.01)) {
  data_new <- data %>% mutate(estimate = ifelse(fit >= p, 1, 0))
  cost <- 1200 * sum(data_new$estimate)
  save <- (sum(data_new$readmit30 + data_new$estimate == 2)) * 0.4 * 8000
  net <- save - cost
  n = n + 1
  outcome2[n, 1] = p
  outcome2[n, 2] = net
}

plot(outcome2[,1], outcome2[,2], type = "l", xlab = "p*", ylab = "savings")

max_saving2 <- max(outcome2[,2])
max_saving2

best_s <- outcome2[,1][which(outcome2[,2] == max_saving2)]
best_s
```

From the graph and outcome matrix, we can find out that the best value for the threshold p* is 0.4, under this circumstance, hospital can obtain a cost savings of $495,200 over the status quo.
